Bandit Multiclass Linear Classification: \\ Efficient Algorithms for the Separable Case

### 2019-02-05 arxiv/appendix-multiclass-perceptron.tex
added arxiv folder

Two results are folklore. The first result is
\autoref{theorem:mutliclass-perceptron-mistake-upper-bound} which states that if
examples are linearly separable with margin $\gamma$ and examples have norm
at most $R$ then the algorithm makes at most $\lfloor 2 (R/\gamma)^2 \rfloor$
mistakes. The second result is
\autoref{theorem:online-multiclass-classification-mistake-lower-bound} which
states that under the same assumptions as in
\autoref{theorem:online-multiclass-classification-mistake-lower-bound}
\emph{any} deterministic algorithm for \textsc{Online Multiclass Classification}
must make at least $\lfloor (R/\gamma)^2 \rfloor$ mistakes in the worst case.

\begin{protocol}[h]
\caption{\textsc{Online Multiclass Classification}
\label{algorithm:mutliclass-classification}}
\textbf{Require:} Number of classes $K$, number of rounds $T$. \\
\textbf{Require:} Inner product space $(V,\ip{\cdot}{\cdot})$. \\
\For{$t=1,2,\dots,T$}{
Adversary chooses example $(x_t, y_t) \in V \times \{1,2,\dots,K\}$, where $x_t$ is revealed to the learner.\\
Predict class label $\widehat y_t \in \{1,2,\dots,K\}$.\\
Observe feedback $y_t$
}
\end{protocol}

@@

\begin{theorem}[Mistake upper bound~\cite{Crammer-Singer-2003}]
\label{theorem:mutliclass-perceptron-mistake-upper-bound}
Let $(V, \ip{\cdot}{\cdot})$ be an inner product space, let $K$ be a positive
integer, let $\gamma$ be a positive real number and let $R$ be a non-negative real
number. If $(x_1, y_1), (x_2, y_2), \dots, (x_T, y_T)$ is a sequence of labeled
examples in $V \times \{1,2,\dots,K\}$ that are weakly linearly separable with margin
$\gamma$ and $\norm{x_1}, \norm{x_2}, \dots, \norm{x_T} \le R$
then \textsc{Multiclass Perceptron} algorithm makes at most $\lfloor
2(R/\gamma)^2 \rfloor$ mistakes.
\end{theorem}

\begin{proof}
Let $M = \sum_{t=1}^T \indicator{\widehat y_t \neq y_t}$ be the number of
mistakes the algorithm makes. Since the $K$-tuple $(w_1^{(t)}, w_2^{(t)}, \dots,
w_K^{(t)})$ changes only if a mistake is made, we can upper bound $\sum_{i=1}^K
\norm{w_i^{(t)}}^2$ in terms of number of mistakes.
If a mistake happens in round $t$ then
\begin{align*}
& \sum_{i=1}^K \norm{w_i^{(t+1)}}^2 \\
& = \left(\sum_{i \in \{1,2,\dots,K\} \setminus \{y_t, \widehat y_t\} } \norm{w_i^{(t)}}^2 \right) \\
& \qquad + \norm{w_{y_t}^{(t)} + x_t}^2 + \norm{w_{\widehat y_t}^{(t)} - x_t}^2 \\
& = \left(\sum_{i \in \{1,2,\dots,K\} \setminus \{y_t, \widehat y_t\} } \norm{w_i^{(t)}}^2 \right) + \norm{w_{y_t}^{(t)}}^2 + \norm{w_{\widehat y_t}^{(t)}}^2 \\
& \qquad + 2 \norm{x_t}^2 + 2 \ip{w_{y_t}^{(t)} - w_{\widehat y_t}^{(t)}}{x_t} \\
& = \left(\sum_{i=1}^K \norm{w_i^{(t)}}^2 \right) + 2 \norm{x_t}^2 + 2 \ip{w_{y_t}^{(t)} - w_{\widehat y_t}^{(t)}}{x_t} \\
& \le \left(\sum_{i=1}^K \norm{w_i^{(t)}}^2 \right) + 2 \norm{x_t}^2 \\
& \le \left(\sum_{i=1}^K \norm{w_i^{(t)}}^2 \right) + 2 R^2 \; .
\end{align*}
So each time a mistake happens, $\sum_{i=1}^K \norm{w_i^{(t)}}^2$ increases by at most $2R^2$. Thus,
$$
\sum_{i=1}^K \norm{w_i^{(T+1)}}^2 \le 2R^2 M \; .
$$
Let $w_1^*, w_2^*, \dots, w_K^* \in V$ be vectors satisfying the
\eqref{equation:weak-linear-separability-1} and
\eqref{equation:weak-linear-separability-2}. We lower bound $\sum_{i=1}^K \ip{w_i^*}{w_i^{(t)}}$. This quantity changes
only when a mistakes happens. If mistake happens in round $t$, we have
\begingroup
\allowdisplaybreaks
\begin{align*}
& \sum_{i=1}^K \ip{w_i^*}{w_i^{(t+1)}} \\
& = \left( \sum_{i \in \{1,2,\dots,K\} \setminus \{y_t, \widehat y_t\}} \ip{w_i^*}{w_i^{(t)}} \right) \\
& \qquad + \ip{w_{y_t}^*}{w_{y_t}^{(t)} + x_t} + \ip{w_{\widehat y_t}^*}{w_{\widehat y_t}^{(t)} - x_t} \\
& = \left( \sum_{i=1}^K \ip{w_i^*}{w_i^{(t)}} \right) + \ip{w_{y_t}^* - w_{\widehat y_t}^*}{x_t} \\
& \ge  \left( \sum_{i=1}^K \ip{w_i^*}{w_i^{(t)}} \right) + \gamma \; .
\end{align*}
\endgroup
Thus, after $M$ mistakes,
$$
\sum_{i=1}^K \ip{w_i^*}{w_i^{(T+1)}} \ge \gamma M \; .
$$
We upper bound the left hand side by using Cauchy-Schwartz inequality twice and
the condition \eqref{equation:weak-linear-separability-1} on $w_1^*, w_2^*, \dots,
w_K^*$. We have
\begin{align*}
\sum_{i=1}^K \ip{w_i^*}{w_i^{(T+1)}}
& \le \sum_{i=1}^K \norm{w_i^*} \cdot \norm{w_i^{(T+1)}} \\
& \le \sqrt{\sum_{i=1}^K \norm{w_i^*}^2} \sqrt{\sum_{i=1}^K \norm{w_i^{(T+1)}}^2} \\
& \le \sqrt{\sum_{i=1}^K \norm{w_i^{(T+1)}}^2} \; .
\end{align*}
Combining all inequalities, we get
$$
(\gamma M)^2 \le \sum_{i=1}^K \norm{w_i^{(T+1)}}^2 \le 2R^2 M \; .
$$
We conclude that $M \le 2(R/\gamma)^2$. Since $M$ is an integer, $M \le \lfloor 2(R/\gamma)^2 \rfloor$.
\end{proof}

@@

\begin{proof}
Let $T = \lfloor (R/\gamma)^2 \rfloor$, $V = \R^T$, and for all $t$ in
$\cbr{1,\ldots,T}$, define instance $x_t = R e_t$ where $e_t$ is $t$-th element
of the standard orthonormal basis of $\R^T$.
Let labels $y_1, \ldots, y_T$ be chosen i.i.d uniformly at random from
$\cbr{1,2,\ldots,K}$ and independently of any randomness used by the algirithm
$\calA$.

@@

We now give a lower bound on the number of mistakes $\calA$ makes.
As $y_t$ is chosen uniformly from $\{1,2,\dots,K\}$ and independent of
$\calA$'s randomization
$$
\Exp[ \indicator{\widehat y_t \neq y_t} ] \ge 1 - \frac{1}{K} \ge \frac{1}{2} \; .
$$
Summing over all $t$ in $\cbr{1,\ldots,T}$, we conclude that
$$
\Exp \sbr{ \sum_{t=1}^T \indicator{\widehat y_t \neq y_t} } \geq \frac T 2 = \frac 1 2 \lfloor (R/\gamma)^2 \rfloor,
$$
which completes the proof.
\end{proof}

>>>>

Two results are folklore. The first result is
\autoref{theorem:multiclass-perceptron-mistake-upper-bound} which states that if
examples are linearly separable with margin $\gamma$ and examples have norm
at most $R$ then the algorithm makes at most $\lfloor 2 (R/\gamma)^2 \rfloor$
mistakes. The second result is
\autoref{theorem:online-multiclass-classification-mistake-lower-bound} which
states that under the same assumptions as in
\autoref{theorem:online-multiclass-classification-mistake-lower-bound}
\emph{any} deterministic algorithm for \textsc{Online Multiclass Classification}
must make at least $\lfloor (R/\gamma)^2 \rfloor$ mistakes in the worst case.

\begin{protocol}[h]
\caption{\textsc{Online Multiclass Classification}
\label{algorithm:mutliclass-classification}}
\textbf{Require:} Number of classes $K$, number of rounds $T$. \\
\textbf{Require:} Inner product space $(V,\ip{\cdot}{\cdot})$. \\
\For{$t=1,2,\dots,T$}{
Adversary chooses example $(x_t, y_t) \in V \times \{1,2,\dots,K\}$, where $x_t$ is revealed to the learner.\\
Predict class label $\widehat y_t \in \{1,2,\dots,K\}$.\\
Observe feedback $y_t$.
}
\end{protocol}

@@

\begin{theorem}[Mistake upper bound~\cite{Crammer-Singer-2003}]
\label{theorem:multiclass-perceptron-mistake-upper-bound}
Let $(V, \ip{\cdot}{\cdot})$ be an inner product space, let $K$ be a positive
integer, let $\gamma$ be a positive real number and let $R$ be a non-negative real
number. If $(x_1, y_1), (x_2, y_2), \dots, (x_T, y_T)$ is a sequence of labeled
examples in $V \times \{1,2,\dots,K\}$ that are weakly linearly separable with margin
$\gamma$ and $\norm{x_1}, \norm{x_2}, \dots, \norm{x_T} \le R$
then \textsc{Multiclass Perceptron} algorithm makes at most $\lfloor
2(R/\gamma)^2 \rfloor$ mistakes.
\end{theorem}

\begin{proof}
Let $M = \sum_{t=1}^T \indicator{\widehat y_t \neq y_t}$ be the number of
mistakes the algorithm makes. Since the $K$-tuple $(w_1^{(t)}, w_2^{(t)}, \dots,
w_K^{(t)})$ changes only if a mistake is made, we can upper bound $\sum_{i=1}^K
\norm{w_i^{(t)}}^2$ in terms of number of mistakes.
If a mistake happens in round $t$ then
\begin{align*}
& \sum_{i=1}^K \norm{w_i^{(t+1)}}^2 \\
& = \left(\sum_{i \in \{1,2,\dots,K\} \setminus \{y_t, \widehat y_t\} } \norm{w_i^{(t)}}^2 \right) \\
& \qquad + \norm{w_{y_t}^{(t)} + x_t}^2 + \norm{w_{\widehat y_t}^{(t)} - x_t}^2 \\
& = \left(\sum_{i \in \{1,2,\dots,K\} \setminus \{y_t, \widehat y_t\} } \norm{w_i^{(t)}}^2 \right) + \norm{w_{y_t}^{(t)}}^2 + \norm{w_{\widehat y_t}^{(t)}}^2 \\
& \qquad + 2 \norm{x_t}^2 + 2 \ip{w_{y_t}^{(t)} - w_{\widehat y_t}^{(t)}}{x_t} \\
& = \left(\sum_{i=1}^K \norm{w_i^{(t)}}^2 \right) + 2 \norm{x_t}^2 + 2 \ip{w_{y_t}^{(t)} - w_{\widehat y_t}^{(t)}}{x_t} \\
& \le \left(\sum_{i=1}^K \norm{w_i^{(t)}}^2 \right) + 2 \norm{x_t}^2 \\
& \le \left(\sum_{i=1}^K \norm{w_i^{(t)}}^2 \right) + 2 R^2 \; .
\end{align*}
So each time a mistake happens, $\sum_{i=1}^K \norm{w_i^{(t)}}^2$ increases by at most $2R^2$. Thus,
\begin{equation}
\sum_{i=1}^K \norm{w_i^{(T+1)}}^2 \le 2R^2 M \; .
\label{equation:perceptron-norm-ub}
\end{equation}
Let $w_1^*, w_2^*, \dots, w_K^* \in V$ be vectors satisfying
\eqref{equation:weak-linear-separability-1} and
\eqref{equation:weak-linear-separability-2}. We lower bound $\sum_{i=1}^K \ip{w_i^*}{w_i^{(t)}}$. This quantity changes
only when a mistakes happens. If mistake happens in round $t$, we have
\begingroup
\allowdisplaybreaks
\begin{align*}
& \sum_{i=1}^K \ip{w_i^*}{w_i^{(t+1)}} \\
& = \left( \sum_{i \in \{1,2,\dots,K\} \setminus \{y_t, \widehat y_t\}} \ip{w_i^*}{w_i^{(t)}} \right) \\
& \qquad + \ip{w_{y_t}^*}{w_{y_t}^{(t)} + x_t} + \ip{w_{\widehat y_t}^*}{w_{\widehat y_t}^{(t)} - x_t} \\
& = \left( \sum_{i=1}^K \ip{w_i^*}{w_i^{(t)}} \right) + \ip{w_{y_t}^* - w_{\widehat y_t}^*}{x_t} \\
& \ge  \left( \sum_{i=1}^K \ip{w_i^*}{w_i^{(t)}} \right) + \gamma \; .
\end{align*}
\endgroup
Thus, after $M$ mistakes,
\[
\label{equation:perceptron-ip-lb}
\sum_{i=1}^K \ip{w_i^*}{w_i^{(T+1)}} \ge \gamma M \; .
\]
We upper bound the left hand side by using Cauchy-Schwartz inequality twice and
the condition \eqref{equation:weak-linear-separability-1} on $w_1^*, w_2^*, \dots,
w_K^*$. We have
\begin{align*}
\sum_{i=1}^K \ip{w_i^*}{w_i^{(T+1)}}
& \le \sum_{i=1}^K \norm{w_i^*} \cdot \norm{w_i^{(T+1)}} \\
& \le \sqrt{\sum_{i=1}^K \norm{w_i^*}^2} \sqrt{\sum_{i=1}^K \norm{w_i^{(T+1)}}^2} \\
& \le \sqrt{\sum_{i=1}^K \norm{w_i^{(T+1)}}^2} \; .
\end{align*}
Combining the above inequality with Equations~\eqref{equation:perceptron-norm-ub}
and~\eqref{equation:perceptron-ip-lb}, we get
$$
(\gamma M)^2 \le \sum_{i=1}^K \norm{w_i^{(T+1)}}^2 \le 2R^2 M \; .
$$
We conclude that $M \le 2(R/\gamma)^2$. Since $M$ is an integer, $M \le \lfloor 2(R/\gamma)^2 \rfloor$.
\end{proof}

@@

\begin{proof}
Let $T = \lfloor (R/\gamma)^2 \rfloor$, $V = \R^T$, and for all $t$ in
$\cbr{1,\ldots,T}$, define instance $x_t = R e_t$ where $e_t$ is $t$-th element
of the standard orthonormal basis of $\R^T$.
Let labels $y_1, \ldots, y_T$ be chosen i.i.d uniformly at random from
$\cbr{1,2,\ldots,K}$ and independently of any randomness used by the algorithm
$\calA$.

@@

We now give a lower bound on the number of mistakes $\calA$ makes.
As $y_t$ is chosen uniformly from $\{1,2,\dots,K\}$, independently from
$\calA$'s randomization and the first $t-1$ examples,
$$
\Exp[ \indicator{\widehat y_t \neq y_t} ] \ge 1 - \frac{1}{K} \ge \frac{1}{2} \; .
$$
Summing over all $t$ in $\cbr{1,\ldots,T}$, we conclude that
$$
\Exp \sbr{ \sum_{t=1}^T \indicator{\widehat y_t \neq y_t} } \geq \frac T 2 = \frac 1 2 \lfloor (R/\gamma)^2 \rfloor,
$$
which completes the proof.
\end{proof}

### 2019-02-05 arxiv/appendix-nearest-neighbor-algorithm.tex
added arxiv folder

\begin{algorithm}[H]
\SetAlgoLined
\LinesNumbered
\caption{\textsc{Nearest-Neighbor Algorithm}
\label{algorithm:nearest-neighbor}
}
\textbf{Require:} Number of classes $K$, number of rounds $T$. \\
\textbf{Require:} Inner product space $(V,\ip{\cdot}{\cdot})$. \\
\nl Initialize $S \gets \emptyset$ \\
\nl \For{$t=1,2,\ldots,T$:}{
\nl  \If{$\min_{(x,y) \in S} \norm{x_t - x} \le \gamma$}{
\nl    Find nearest neighbor \\
       \qquad \qquad $(\widetilde{x}, \widetilde{y}) = \argmin_{(x,y) \in S} \norm{x_t - x}$ \\
\nl	  Predict $\widehat{y}_t = \widetilde{y}$
  }
  \Else{
\nl	  Predict $\widehat y_t \sim \text{Uniform}(\{1,2,\dots,K\})$
      \label{line:nearest-neighbor-explore} \\
\nl	  Receive feedback $z_t = \indicator{\widehat{y}_t \neq y_t}$ \\
\nl	  \If{$z_t = 0$}{
\nl      $S \gets S \cup \cbr{(x_t, \widehat{y}_t)}$
    }
  }
}
\end{algorithm}

In this section we analyze \textsc{Nearest-Neighbor Algorithm} shown as
Algorithm~\ref{algorithm:nearest-neighbor}. The algorithm is based on obvious
the idea that under weak linear assumption two examples that are close to each
other must have the same label. Lemma below formalizes this intuition.

@@

We also need to define several notions. A subset $S \subseteq \R^d$ is called a
$\gamma$-packing if for any $x,x' \in S$ such that $x \neq x'$ we have $\norm{x -
x'} > \gamma$. The following lemma is standard. Also recall that $\B(x,R) = \{ x'
\in \R^d ~:~ \norm{x' - x} \le R \}$ denotes the closed ball of radius $R$
centered a point $x$.

>>>>

\begin{algorithm}[H]
\SetAlgoLined
\LinesNumbered
\setcounter{AlgoLine}{0}
\caption{\textsc{Nearest-Neighbor Algorithm}
\label{algorithm:nearest-neighbor}
}
\textbf{Require:} Number of classes $K$, number of rounds $T$. \\
\textbf{Require:} Inner product space $(V,\ip{\cdot}{\cdot})$. \\
\nl Initialize $S \gets \emptyset$ \\
\nl \For{$t=1,2,\ldots,T$:}{
\nl  \If{$\min_{(x,y) \in S} \norm{x_t - x} \le \gamma$}{
\nl    Find nearest neighbor \\
       \qquad \qquad $(\widetilde{x}, \widetilde{y}) = \argmin_{(x,y) \in S} \norm{x_t - x}$ \\
\nl	  Predict $\widehat{y}_t = \widetilde{y}$
  }
\nl \Else{
\nl	  Predict $\widehat y_t \sim \text{Uniform}(\{1,2,\dots,K\})$
      \label{line:nearest-neighbor-explore} \\
\nl	  Receive feedback $z_t = \indicator{\widehat{y}_t \neq y_t}$ \\
\nl	  \If{$z_t = 0$}{
\nl      $S \gets S \cup \cbr{(x_t, \widehat{y}_t)}$
    }
  }
}
\end{algorithm}

In this section we analyze \textsc{Nearest-Neighbor Algorithm} shown as
Algorithm~\ref{algorithm:nearest-neighbor}. The algorithm is based on the obvious
idea that, under the weak linear separability assumption, two examples that are close to each
other must have the same label. The lemma below formalizes this intuition.

@@

We also need to define several notions. A subset $S \subseteq \R^d$ is called a
$\gamma$-packing if for any $x,x' \in S$ such that $x \neq x'$ we have $\norm{x -
x'} > \gamma$. The following lemma is standard. Also recall that $\B(x,R) = \{ x'
\in \R^d ~:~ \norm{x' - x} \le R \}$ denotes the closed ball of radius $R$
centered a point $x$. For set $S \subseteq \R^d$, denote by $\Vol(S)$ the volume of
$S$.

### 2019-02-05 arxiv/appendix-proofs-for-strongly-separable-examples.tex
added arxiv folder

To finish the proof, we need to upper bound the number of updates $U$. We claim
that $U \le \lfloor 4(R/\gamma)^2 \rfloor$ with probability 1.
The proof of this upper bound is
similar to the proof of the mistake bound for \textsc{Multiclass Perceptron}
algorithm. Let $w_1^*, w_2^*, \dots, w_K^* \in V$ be vectors that satisfy
\eqref{equation:strong-linear-separability-1},
\eqref{equation:strong-linear-separability-2} and
\eqref{equation:strong-linear-separability-3}.
The $K$-tuple $(w_1^{(t)}, w_2^{(t)}, \dots, w_K^{(t)})$
changes only if there is an update in round $t$.
We investigate how $\sum_{i=1}^K \norm{w_i^{(t)}}^2$ and
$\sum_{i=1}^K \ip{w_i^*}{w_i^{(t)}}$ change. If there is an update in round $t$,
by lines~\ref{line:pos-update} and~\ref{line:neg-update}, we always have
$ w_{\widehat y_t}^{(t+1)} = w_{\widehat y_t}^{(t)} + (-1)^{z_t} x_t $,
and for all $i \neq \widehat y_t$, $w_{i}^{(t+1)} = w_{i}^{(t)}$.
Therefore,
\begingroup
\allowdisplaybreaks
\begin{align*}
& \sum_{i=1}^K \norm{w_i^{(t+1)}}^2 \\
& = \left( \sum_{i \in \{1,2,\dots,K\} \setminus \{\widehat y_t\}} \norm{w_i^{(t)}}^2 \right) + \norm{w_{\widehat y_t}^{(t+1)}}^2 \\
& = \left( \sum_{i \in \{1,2,\dots,K\} \setminus \{\widehat y_t\}} \norm{w_i^{(t)}}^2 \right) + \norm{w_{\widehat y_t}^{(t)} + (-1)^{z_t} x_t}^2 \\
& = \left( \sum_{i=1}^K \norm{w_i^{(t)}}^2 \right) + \norm{x_t}^2 + \underbrace{(-1)^{z_t} 2 \ip{w_{\widehat y_t}^{(t)}}{x_t}}_{\le 0} \\
& \le \left( \sum_{i=1}^K \norm{w_i^{(t)}}^2 \right) + \norm{x_t}^2 \\
& \le \left( \sum_{i=1}^K \norm{w_i^{(t)}}^2 \right) + R^2 \; .
\end{align*}
\endgroup
The inequality that $(-1)^{z_t} 2 \ip{w_{\widehat y_t}^{(t)}}{x_t} \leq 0$ is from a case analysis: if line~\ref{line:pos-update} is executed, then $z_t = 0$ and $\ip{w_i^{(t)}}{x_t} \ge 0$; otherwise line~\ref{line:neg-update} is executed, in which case $z_t = 1$ and $\ip{w_i^{(t)}}{x_t} \le 0$.

@@

The norm of each example is exactly $R$. The examples are strongly separable
with margin $\gamma$. To see that, consider $w_1^*, w_2^*, \dots, w_K^* \in V$
defined by
$$
w_i^* = \sqrt{2} \frac{\gamma}{R} \left( \sum_{j ~:~ \ell_j = i} e_j \right) - \frac{\sqrt{2}}2 \frac{\gamma}{R} e_{M+1}
$$
for $i=1,2,\dots,K$.

>>>>

To finish the proof, we need to upper bound the number of updates $U$. We claim
that $U \le \lfloor 4(R/\gamma)^2 \rfloor$ with probability 1.
The proof of this upper bound is
similar to the proof of the mistake bound for \textsc{Multiclass Perceptron}
algorithm. Let $w_1^*, w_2^*, \dots, w_K^* \in V$ be vectors that satisfy
\eqref{equation:strong-linear-separability-1},
\eqref{equation:strong-linear-separability-2} and
\eqref{equation:strong-linear-separability-3}.
The $K$-tuple $(w_1^{(t)}, w_2^{(t)}, \dots, w_K^{(t)})$
changes only if there is an update in round $t$.
We investigate how $\sum_{i=1}^K \norm{w_i^{(t)}}^2$ and
$\sum_{i=1}^K \ip{w_i^*}{w_i^{(t)}}$ change. If there is an update in round $t$,
by lines~\ref{line:pos-update} and~\ref{line:neg-update}, we always have
$ w_{\widehat y_t}^{(t+1)} = w_{\widehat y_t}^{(t)} + (-1)^{z_t} x_t $,
and for all $i \neq \widehat y_t$, $w_{i}^{(t+1)} = w_{i}^{(t)}$.
Therefore,
\begingroup
\allowdisplaybreaks
\begin{align*}
& \sum_{i=1}^K \norm{w_i^{(t+1)}}^2 \\
& = \left( \sum_{i \in \{1,2,\dots,K\} \setminus \{\widehat y_t\}} \norm{w_i^{(t)}}^2 \right) + \norm{w_{\widehat y_t}^{(t+1)}}^2 \\
& = \left( \sum_{i \in \{1,2,\dots,K\} \setminus \{\widehat y_t\}} \norm{w_i^{(t)}}^2 \right) + \norm{w_{\widehat y_t}^{(t)} + (-1)^{z_t} x_t}^2 \\
& = \left( \sum_{i=1}^K \norm{w_i^{(t)}}^2 \right) + \norm{x_t}^2 + \underbrace{(-1)^{z_t} 2 \ip{w_{\widehat y_t}^{(t)}}{x_t}}_{\le 0} \\
& \le \left( \sum_{i=1}^K \norm{w_i^{(t)}}^2 \right) + \norm{x_t}^2 \\
& \le \left( \sum_{i=1}^K \norm{w_i^{(t)}}^2 \right) + R^2 \; .
\end{align*}
\endgroup
The inequality that $(-1)^{z_t} 2 \ip{w_{\widehat y_t}^{(t)}}{x_t} \leq 0$ is from a case analysis: if line~\ref{line:pos-update} is executed, then $z_t = 0$ and $\ip{w_{\widehat y_t}^{(t)}}{x_t} < 0$;
 otherwise line~\ref{line:neg-update} is executed, in which case $z_t = 1$ and $\ip{w_{\widehat y_t}^{(t)}}{x_t} \ge 0$.

@@

The norm of each example is exactly $R$. The examples are strongly linearly separable
with margin $\gamma$. To see that, consider $w_1^*, w_2^*, \dots, w_K^* \in V$
defined by
$$
w_i^* = \sqrt{2} \frac{\gamma}{R} \left( \sum_{j ~:~ \ell_j = i} e_j \right) - \frac{\sqrt{2}}2 \frac{\gamma}{R} e_{M+1}
$$
for $i=1,2,\dots,K$.

### 2019-02-05 arxiv/empirical-comparison.tex
added arxiv folder

We implemented
Algorithm~\ref{algorithm:algorithm-for-strongly-linearly-separable-examples},
Algorithm~\ref{algorithm:kernelized} with rational
kernel~\eqref{equation:rational-kernel} and used implementation of
\textsc{Banditron} algorithm by \citet{Orabona09}.\footnote{All the source code
is included in the supplementary material.}
We evaluated these algorithms on the two datasets. \textsc{Banditron} has an
exploration rate parameter, for which we tried values $0.02, 0.01, 0.005, 0.002,
0.001, 0.0005$. Since all three algorithms are randomized, we run each algorithm
$20$ times. The average cumulative number of mistakes up to round $t$ as a
function of $t$ are shown in
Figures~\ref{figure:number-of-mistakes-strongly-separable-dataset}
and~\ref{figure:number-of-mistakes-weakly-separable-dataset}.

We can see that there is a tradeoff in the setting of the exploration rate for
\textsc{Banditron}. With large exploration parameter, \textsc{Banditron} suffers
from over-exploration, whereas with small exploration parameter, its model
cannot be updated quickly enough. As expected,
Algorithm~\ref{algorithm:algorithm-for-strongly-linearly-separable-examples} has
a small number of mistakes in the strongly linearly separable setting, while
having a large number of mistakes in the weakly linearly separable setting, due
to the limited representation power of linear classifiers. In contrast,
Algorithm~\ref{algorithm:kernelized} with rational kernel has a small number of
mistakes in both settings, exhibiting strong adaptivity guarantees.
Appendix~\ref{section:supp-to-experiment} shows the decision boundaries that
each of the algorithms learns.

>>>>

We implemented
Algorithm~\ref{algorithm:algorithm-for-strongly-linearly-separable-examples},
Algorithm~\ref{algorithm:kernelized} with rational
kernel~\eqref{equation:rational-kernel} and used implementation of
\textsc{Banditron} algorithm by \citet{Orabona09}.\footnote{The source code
is included in the supplementary material.}
We evaluated these algorithms on the two datasets. \textsc{Banditron} has an
exploration rate parameter, for which we tried values $0.02, 0.01, 0.005, 0.002,
0.001, 0.0005$. Since all three algorithms are randomized, we run each algorithm
$20$ times. The average cumulative number of mistakes up to round $t$ as a
function of $t$ are shown in
Figures~\ref{figure:number-of-mistakes-strongly-separable-dataset}
and~\ref{figure:number-of-mistakes-weakly-separable-dataset}.

We can see that there is a tradeoff in the setting of the exploration rate for
\textsc{Banditron}. With large exploration parameter, \textsc{Banditron} suffers
from over-exploration, whereas with small exploration parameter, its model
cannot be updated quickly enough. As expected,
Algorithm~\ref{algorithm:algorithm-for-strongly-linearly-separable-examples} has
a small number of mistakes in the strongly linearly separable setting, while
having a large number of mistakes in the weakly linearly separable setting, due
to the limited representation power of linear classifiers. In contrast,
Algorithm~\ref{algorithm:kernelized} with rational kernel has a small number of
mistakes in both settings, exhibiting strong adaptivity guarantees.
Appendix~\ref{section:supp-to-experiment} shows the decision boundaries that
each of the algorithms learns by the end of the last round.
